\subsection{Deployed Software \& Versions}
To sum up, the cluster has been provisioned with HDP 2.6.1, which includes:
\begin{itemize}
    \item Hadoop 2.7.3.2.6.1.0-129 (HDFS, YARN and MapReduce),
    \item Tez 0.7.0,
    \item Hive 1.2.1000, 
    \item ZooKeeper 3.4.6,
    \item Kafka 0.10.1,
    \item Knox 0.12.0,
    \item Ranger 0.7.0,
    \item Spark 2.1.1,
    \item NiFi 1.2.0,
    \item Registry 0.3.0,
    \item Slider 0.92.0.
\end{itemize}

In addition, it's been used Flink 1.4.0 for Scala 2.11 compiled from sources against hadoop 2.7.3.2.6.1.0-129 and Cassandra 3.11.
\\
\\
The cluster is composed of seven machines:
\begin{itemize}
    \item \texttt{frontend}, with Knox and a Nginx reverse proxy installed, used as single point of entry for developers and cluster operators to access cluster services via SSH tunneling. Specifications: 1 core, 512 MB of RAM.
    \item \texttt{master-1}, one of the two master components of the cluster, with an HDFS Namenode, a Zookeeper Server, a Registry MySQL database, a Kafka Broker, Hive Metastore and HiveServer 2 Interactive (support for Hive LLAP daemons), Ranger Admin and UserSync, a NiFi node and Spark2 History Server. Specifications: 4 core, 12 GB of RAM.
    \item \texttt{master-2}, the other master component of the cluster, with the YARN Resource Manager and App Timeline Server, HDFS Secondary Namenode, a Zookeeper Server, a WebHCat Server and HiveServer2 for Hive, a Kafka Broker, a Nifi node, MapReduce2 History Server and Flink binaries. Specifications: 4 core, 12 GB of RAM.
    \item Four slave machines, \texttt{slave-1}, \texttt{slave-2}, \texttt{slave-3}, \texttt{slave-4}, with a YARN Node Manager, HDFS DataNode and Cassandra nodes. Specifications: 4 core, 6 GB of RAM.
\end{itemize}